---
title: "SBDI_tree_nets"
author: "Mariana Braga"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
execute:
  warning: false
---


```{r}
#| label: "packages"
 
library(tidyverse)
library(bipartite)
library(tidygraph)
library(ggraph)
```

Read data and count entries by interaction type

```{r}
#| label: "read data"

raw_data <- read.table("data/report_data.txt", sep = "\t", header = T, fileEncoding = "ISO-8859-1")

# check for duplicates
nrow(raw_data) > n_distinct(raw_data[,c("Scientificname", "VärdSciName", "Nyttjande")])
```

There are duplicates in the data (seems to be for different parts of the tree). Remove them from data subsets later.

```{r}
# number of tree host species
n_distinct(raw_data$VärdCommon)
# number of attackers
n_distinct(raw_data$Commonname)

# taxonomic data from the report's appendix 2
tree_taxo <- read.csv("data/tree_taxonomy.csv")

# number of trees
n_distinct(tree_taxo$Scientific.name)

data_trees <- filter(raw_data, VärdSciName %in% tree_taxo$Scientific.name)

# number of attackers by type of interaction 
table(data_trees$Nyttjande) %>% sort()
```

### Parasites

```{r}
data_parasitism <- filter(data_trees, Nyttjande == "Parasitism") %>% 
  select("Scientificname", "VärdSciName", "tal1") %>% 
  rename("parasite" = "Scientificname",
         "host" = "VärdSciName", 
         "weight" = "tal1")

n_distinct(data_parasitism$parasite)
n_distinct(data_parasitism$host)
nrow(data_parasitism)
n_distinct(data_parasitism)

data_parasitism %>% 
  group_by(parasite, host, weight) %>% 
  summarise(n = n()) %>% 
  filter(n > 1) %>% 
  nrow()

data_parasitism %>% 
  group_by(parasite, host) %>% 
  summarise(n = n()) %>% 
  filter(n > 1) %>% 
  nrow()

raw_data %>% 
  filter(Scientificname == "Biorhiza pallida")

raw_data %>% 
  filter(Scientificname == "Andricus quercuscorticis")

data_parasitism <- unique(data_parasitism)

duplicates_para <- data_parasitism %>% 
  group_by(parasite, host) %>% 
  summarise(n = n()) %>% 
  filter(n > 1)

# this interaction is recorded as both 1 and 2
raw_data %>% 
  filter(Scientificname == "Gymnosporangium clavariiforme")

data_parasitism <- data_parasitism[-which(data_parasitism$parasite == duplicates_para$parasite &
                                         data_parasitism$host == duplicates_para$host &
                                         data_parasitism$weight == 1),]
```

# Plot trees x parasite species network

```{r}
# with ggraph

nodes_para <- data.frame(name = c(unique(data_parasitism$parasite), unique(data_parasitism$host)),
                         type = c(rep("parasite",n_distinct(data_parasitism$parasite)), 
                                  rep("host",n_distinct(data_parasitism$host))))

parasit_graph <- tbl_graph(nodes = nodes_para, edges = data_parasitism)

ggraph(parasit_graph, layout = "kk") + 
  geom_edge_link(aes(edge_width = weight), color = "grey30", alpha = 0.8) + 
  geom_node_point(aes(shape = type, color = type), size = 3) +
  scale_color_manual(values = c("chartreuse3","deepskyblue3")) +
  scale_edge_width(range = c(0.7,1.4))
```

```{r}
# with bipartite

adjm_parasitism <- frame2webs(mutate(data_parasitism, webID = 1), varnames = c("host","parasite", "webID", "weight"))[[1]]

plotweb(adjm_parasitism, text.rot = 90, y.lim = c(0, 2))
visweb(adjm_parasitism, prednames = F, preynames = F)
```

```{r}
#| eval: false
# modularity
mod_para <- computeModules(adjm_parasitism)
plotModuleWeb(mod_para)
```


### Herbivores

## Species

```{r}
data_herbivory <- filter(data_trees, Nyttjande == "Föda") %>% 
  select("Scientificname", "VärdSciName", "tal1") %>% 
  rename("herbivore" = "Scientificname",
         "host" = "VärdSciName", 
         "weight" = "tal1")

n_distinct(data_herbivory$herbivore)
n_distinct(data_herbivory$host)
nrow(data_herbivory)
n_distinct(data_herbivory)

duplicates_herb <- data_herbivory %>% 
  group_by(herbivore, host, weight) %>% 
  summarise(n = n()) %>% 
  filter(n > 1)

nrow(duplicates_herb)

raw_data %>% 
  filter(Scientificname == "Aleurocystidiellum disciforme",
         Nyttjande == "Föda")

raw_data %>% 
  filter(Scientificname == "Caloptilia cuculipennella",
         Nyttjande == "Föda")

data_herbivory <- unique(data_herbivory)

duplicates_herb <- data_herbivory %>% 
  group_by(herbivore, host) %>% 
  summarise(n = n()) %>% 
  filter(n > 1)

View(duplicates_herb)

data_herbivory <- data_herbivory[-which(data_herbivory$herbivore == duplicates_herb$herbivore &
                                        data_herbivory$host == duplicates_herb$host &
                                        data_herbivory$weight == 1),]

to_remove <- c()
for(i in 1:nrow(duplicates_herb)) {
  row <- which(data_herbivory$herbivore == duplicates_herb$herbivore[i] &
                              data_herbivory$host == duplicates_herb$host[i] &
                              data_herbivory$weight == 1)
  to_remove <- c(to_remove, row)
}

data_herbivory <- data_herbivory[-to_remove,]
nrow(data_herbivory) == n_distinct(data_herbivory)
```

	
# Plot trees x herbivore species network

```{r}
# with ggraph

nodes_herb <- data.frame(name = c(unique(data_herbivory$herbivore), unique(data_herbivory$host)),
                         type = c(rep("herbivore",n_distinct(data_herbivory$herbivore)), 
                                  rep("host",n_distinct(data_herbivory$host))))

herb_graph <- tbl_graph(nodes = nodes_herb, edges = data_herbivory)

ggraph(herb_graph, layout = "kk") + 
  geom_edge_link(aes(edge_width = weight), color = "grey30", alpha = 0.8) + 
  geom_node_point(aes(shape = type, color = type), size = 2) +
  scale_color_manual(values = c("deepskyblue3","chartreuse3")) +
  scale_edge_width(range = c(0.5,1))
```

```{r}
# with bipartite

adjm_herb <- frame2webs(mutate(data_herbivory, webID = 1), varnames = c("host","herbivore", "webID", "weight"))[[1]]

plotweb(adjm_herb, text.rot = 90, y.lim = c(0, 2), labsize = 0.7)

#visweb(adjm_herb, prednames = F, preynames = F)
```

```{r}
#| eval: false
# modularity
mod_herb <- computeModules(adjm_herb)
plotModuleWeb(mod_herb, displayBlabels = F, labsize = 0.4)

```

## Families

```{r}
data_herb_fam <- filter(data_trees, Nyttjande == "Föda") %>% 
  select("Family", "VärdSciName", "tal1") %>% 
  rename("herb_fam" = "Family",
         "host" = "VärdSciName", 
         "weight" = "tal1")

n_distinct(data_herb_fam$herb_fam)
n_distinct(data_herb_fam$host)
nrow(data_herb_fam)
n_distinct(data_herb_fam)

data_herb_fam <- unique(data_herb_fam)

duplicates_herb_fam <- data_herb_fam %>% 
  group_by(herb_fam, host) %>% 
  summarise(n = n()) %>% 
  filter(n > 1)

# if there are interaction with weight 1 and 2, keep only 2
to_remove_fam <- c()
for(i in 1:nrow(duplicates_herb_fam)) {
  row <- which(data_herb_fam$herb_fam == duplicates_herb_fam$herb_fam[i] &
               data_herb_fam$host == duplicates_herb_fam$host[i] &
               data_herb_fam$weight == 1)
  to_remove_fam <- c(to_remove_fam, row)
}

data_herb_fam <- data_herb_fam[-to_remove_fam,]
n_distinct(data_herb_fam$herb_fam)
n_distinct(data_herb_fam$host)
```

# Plot trees x herbivore families network

```{r}
# with ggraph

nodes_herb_fam <- data.frame(name = c(unique(data_herb_fam$herb_fam), unique(data_herb_fam$host)),
                             type = c(rep("herb_fam",n_distinct(data_herb_fam$herb_fam)), 
                                      rep("host",n_distinct(data_herb_fam$host))))

herb_fam_graph <- tbl_graph(nodes = nodes_herb_fam, edges = data_herb_fam)

ggraph(herb_fam_graph, layout = "auto") + 
  geom_edge_link(aes(edge_width = weight), color = "grey30", alpha = 0.8) + 
  geom_node_point(aes(shape = type, color = type), size = 3) +
  scale_color_manual(values = c("deepskyblue3","chartreuse3")) +
  scale_edge_width(range = c(0.5,1))
```

```{r}
# with bipartite

adjm_herb_fam <- frame2webs(mutate(data_herb_fam, webID = 1), varnames = c("host","herb_fam", "webID", "weight"))[[1]]

plotweb(adjm_herb_fam, text.rot = 90, y.lim = c(0, 2), labsize = 0.7)

visweb(adjm_herb_fam, labsize = 4)
visweb(adjm_herb_fam, labsize = 4, type = "diagonal")
```

```{r}
#| eval: false
# modularity
mod_herb_fam <- computeModules(adjm_herb_fam)
plotModuleWeb(mod_herb_fam, labsize = 0.4)

```

## Families weighted by number of species interactions

```{r}
data_herb_spp_fam <- filter(data_trees, Nyttjande == "Föda") %>% 
  select("Scientificname", "Family", "VärdSciName") %>% 
  rename("herbivore" = "Scientificname",
         "herb_fam" = "Family",
         "host" = "VärdSciName") %>% 
  unique() %>% 
  group_by(herb_fam, host) %>% 
  summarise(n_spp = n()) %>% 
  mutate(logn = round(log(n_spp) + 1))

```

Plot network with log(number of species that interact) + 1 so that low weight interactions are visible.

```{r}
# with bipartite
adjm_herb_spp_fam <- frame2webs(mutate(data_herb_spp_fam, webID = 1), varnames = c("host","herb_fam", "webID", "n_spp"))[[1]]
adjm_herb_spp_fam_log <- frame2webs(mutate(data_herb_spp_fam, webID = 1), varnames = c("host","herb_fam", "webID", "logn"))[[1]]

#plotweb(adjm_herb_spp_fam_log, text.rot = 90, y.lim = c(0, 2), labsize = 0.7)
visweb(adjm_herb_spp_fam_log, labsize = 4)
```

```{r}
#| eval: false
# modularity
mod_herb_spp_fam <- computeModules(adjm_herb_spp_fam_log)
plotModuleWeb(mod_herb_spp_fam, labsize = 0.4)

```


